---
title: "mtcars_Analysis"
author: "Ash Thompson"
date: "16 December 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Regression Models Course Project
## Executive Summary
This report is an analysis of the mtcars dataset, to explore the relationship betweena set of variables and miles per gallon (MPG). 
Particularly interested in answering the following two questions:
- Is an automatic or manual transmission better for MPG
- Quantify the MPG difference between automatic and manual transmissions

## Initial steps
The first steps are to load the dataset and supporting libraries, then perform some minor transformations and exploratory analysis.

```{r initial_steps, message=FALSE}
data(mtcars)
library(ggplot2)
library(dplyr)
head(mtcars)
mtcars$am <- factor(mtcars$am, labels =c("Automatic", "Manual"))
```
Please see appendix 1 for a box plot of the 2 transmission types and a comparison of mpg, that shows just
looking at transmission type, it appears manual transmission gives around 7 more mpg.

## Using linear regression to investigate further
These steps are to investigate the hypothesis that manual transmission gives 7 more 
mpg. First a simple linear model.

```{r model1}
model1 <- lm(mpg ~ am, data=mtcars)
summary(model1)
```

R squared value is fairly low at 0.3385, so only around 34% of the variance can 
be explained by this model. the next steps are to include the other variables into
the model and do an analysis of variance to determine the best variables to include
in a final model.

``` {r model2}
model2 <- lm(mpg ~ ., data=mtcars)
summary(model2)
summary(aov(mpg ~ ., data=mtcars))
```
The variables cyl, disp and wt are significant enough to include in the model, in addition to 
transmission type.

``` {r model3}
model3 <- lm(mpg ~ am + cyl + disp + wt, data=mtcars)
summary(model3)
```
Please see appendix 2 for the plot of residuals showing they have roughly the same variance.
The Normal Q-Q plot shows that the distribution of residuals is normal.

## Conclusion
Manual transmission, taking other variables into account, gives a slightly higher MPG value of approximately 0.1. But number of cylinders and 
weight have more of an effect, cylinders lowering it by around 1.7 and wt lowering by 3.5.

## Appendix 1
```{r appendix1, echo=FALSE}
#summary of manual transmission mpg
print("Manual")
summary(select(filter(mtcars, am=="Manual"),mpg))

#mean of automatic transmission mpg
print("Automatic")
summary(select(filter(mtcars, am=="Automatic"),mpg))

boxplot(mpg ~ am, data = mtcars, col = (c("green","yellow")), ylab = "mpg(Miles Per Gallon)", xlab = "am(Transmission Type)")

```

## Appendix 2

```{r appendix2, echo=FALSE}
par(mfrow = c(2, 2))
plot(model3)

```

